On the effects of partying on next-day sports performance

### 2017-05-29 writing/draft.tex
Added citations toward alcohol and sleep-deprivation. Cleaned up intro to section three on materials and methods.

\usepackage{hyperref}

@@

Studies on effects of alcohol on sports performance in a controlled setting
show that alcohol impairs the nervous system, resulting in decrements in both
cognitive function and motor skill.\cite{shirreffs} In a study on 
the effects of alcohol on recovery of male athletes, it was found that
consumption is correlated with reduced production of testosterone
which ultimately leads to inhibited ability to recover and adapt to exercise.\cite{mjbarnes}
Same goes for sleep deprivation and possibly narcotics. 
These are all well correlated with active night life.

\section{Materials and Methods}
One of the hardest parts of this analysis is that partying is a latent variable.
In order to determine when players partied, one might consider to look at social media data
for direct mentions of partying, either by athletes or witnesses. However, due to
intense public scrutiny that athletes face, there is a strong reluctance to post
such mentions on social media. There are notable exceptions, such
as Oddel Beckham Jr.'s trip to Miami in 2017 season.\cite{bleler}
The reports are few and far enough between,
and could be biased based on whether a particular city has nightclubs with VIP rooms,
that picking up statistical evidence would be fruitless.

\paragraph{Two approaches: a discrete indicator and a continuous measure of nightlife}
Because famous athletes often enjoy a celebrity lifestyle, using conventional statistics
such as number of drinking establishments in each city does not work; athletes instead
tend to visit high end nightclubs. We again turned to player interviews for insight,
and found two major insights. We obtained a listing of cities which were anecdotally
mentioned as having active night-life. Between players, the rank ordering of these
was consistent to the extent that LA and NY were top
offenders. This motivates using a binary indicator for party in these two cities. 
We also found that athletes tend to party with other A-list celebrities,
particularly musicians. This motivated a continuous measure of nightlife, based on a
simple count of the number of musicians. We pursue both approaches.

Insights: major party cities are LA and NY, and athletes party is well correlated with musicians. This led to two separate definitions of party. Our most basic analysis was to consider performance 
the day after visiting LA or NY.
We also considered the number of musical establishments per city.

@@

For example, if book-keepers
are already aware of the effects of night-life on performance, it could
already be factored into the spread; if this were the case, then our variable
of interest would not appear to have significant explanatory power
on the margin, when also conditioning on spread, in spite of there being
a true effect. Alternatively,
if the house book-keeper systematically over-valued
teams which you are more likely to play after a party city,
we could see a spurious effect.
That said, the amount of money that is at stake and the effort put into creating these spreads 
suggests that they are reliable. So, if we do find an effect in our variable of interest,
we can be confident in our conclusion that there is a causal effect and not some
other systematic bias.

>>>>

\usepackage{hyperref}
\usepackage{color,soul}

@@

Studies on effects of alcohol on sports performance in a controlled setting
show that alcohol impairs the nervous system, resulting in decrements in both
cognitive function and motor skill.\cite{shirreffs} In a study on 
the effects of alcohol on recovery of male athletes, it was found that
consumption is correlated with reduced production of testosterone
which ultimately leads to inhibited ability to recover and adapt to exercise.\cite{mjbarnes}
Sleep deprivation has also been shown to strongly impair cognitive and motor performance.\cite{pilcher} Both alcohol and sleep deprivation are well correlated with lifestyles associated
with an active night life.

\section{Materials and Methods}
One of the hardest parts of this analysis is that partying is a latent variable.
In order to determine when players partied, one might consider to look at social media data
for direct mentions of partying, either by athletes or witnesses. However, due to
intense public scrutiny that athletes face, there is a strong reluctance to post
such mentions on social media. There are notable exceptions, such
as Oddel Beckham Jr.'s trip to Miami in 2017 season.\cite{bleler}
The reports are few and far enough between,
and whether they are even reported at all is strongly biased on whether a particular city 
has nightclubs with VIP rooms, etc; picking up statistical evidence would be fruitless in this case.

\paragraph{Two approaches: a discrete indicator and a continuous measure of nightlife}
Because famous athletes often enjoy a celebrity lifestyle, using conventional statistics
such as number of drinking establishments in each city does not work; athletes instead
tend to visit high end nightclubs. We again turned to player interviews for ideas,
and found two major insights. From our conversations, 
we obtained a listing of cities which were anecdotally
mentioned as having active night-life. Between players, the rank ordering of these
was consistent to the extent that LA and NY were top
offenders. This motivates using a binary indicator for party in these two cities. 
We also found that athletes tend to party with other A-list celebrities,
particularly musicians. This motivated a continuous measure of nightlife, based on a
simple count of the number of musicians in the county in which the stadium resides. 
We pursue both approaches.

@@

For example, if book-keepers
are already aware of the effects of night-life on performance, it could
already be factored into the spread; if this were the case, then our variable
of interest would not appear to have significant explanatory power
after conditioning on spread, in spite of there being
a true effect. Alternatively,
if the house book-keeper systematically over-valued
teams which you are more likely to play after a party city,
we could see a spurious effect.

\hl{Careful} That said, the amount of money that is at stake and the effort put into creating these spreads 
suggests that they are reliable. So, if we do find an effect in our variable of interest,
we can be confident in our conclusion that there is a causal effect and not some
other systematic bias.
